```{r setup, echo = FALSE, results = "hide", warning = FALSE, message = FALSE}
require(knitr)
opts_chunk$set(echo = FALSE)
opts_chunk$set(warning = FALSE)
opts_chunk$set(results = "hide")
opts_chunk$set(fig.width = 10)
opts_chunk$set(fig.height = 7)
# knit_hooks$set(inline = knitr:::.inline.hook.tex)
# knitr::render_latex()
library(xtable)
```
```{r loaddata}
# TODO: better strategy: normalize by saturated model cost?
library(lars)
library(cosso)
data(diabetes)
x <- scale(diabetes$x)
n <- nrow(x)
p <- ncol(x)
y <- diabetes$y
maxk <- 50
eps <- 1e-9
best <- order(abs(lm(y ~ x - 1)$coef))
worst <- order(-abs(lm(y ~ x - 1)$coef))

set.seed(95)
train <- sample(1:n, round(n * 0.8), replace = FALSE)
test <- setdiff(1:n, train)
xtrain <- x[train, ]
ytrain <- y[train]
xtrain <- scale(xtrain)
ytrain <- ytrain - mean(ytrain)
xtest <- x[test, ]
ytest <- y[test]
ofgmax <- 1
```

Top row is updated algorithm.

Bottom row is old algorithm.

Left plots are for training data.

Right plots are for test data.

Example 1: all variables costs 1
================================

clars algorithm should be equivalent to lasso


```{r equal}
par(mfrow = c(1, 2))
###########################################################################
##                      all variables cost the same                      ##
###########################################################################


set.seed(95)
cost <- rep(1, p)

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 50, trace = TRUE, ofgmax=ofgmax)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


# set.seed(95)
# cost <- rep(1, p)

# clarsP <- clars(xtrain, ytrain, cost, maxk = 50, trace = TRUE)
# larsP <- lars(xtrain, ytrain, type = "lar")
# lassP <- lars(xtrain, ytrain, type = "lasso")

# clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
# larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
# lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

# xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
#           max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
# ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
#           max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

# plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

# clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
# larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
# lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

# xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
#           max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
# ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
#           max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

# plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))
```

\newpage

Example 2: random cost, same order of magnitude
===============================================

```{r random}

par(mfrow = c(1, 2))

###########################################################################
##                  random cost same order of magnitude                  ##
###########################################################################

set.seed(95)
cost <- runif(p, 1, 10)

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 50, trace = FALSE, ofgmax=ofgmax)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


# set.seed(95)
# cost <- runif(p, 1, 10)

# clarsP <- clars(xtrain, ytrain, cost, maxk = 50, trace = FALSE)
# larsP <- lars(xtrain, ytrain, type = "lar")
# lassP <- lars(xtrain, ytrain, type = "lasso")

# clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
# larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
# lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

# xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
#           max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
# ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
#           max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

# plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

# clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
# larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
# lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

# xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
#           max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
# ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
#           max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

# plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

```

\newpage

Example 3: random cost, different order of magnitude
====================================================

```{r randommag}

par(mfrow = c(1, 2))

###########################################################################
##                random cost different order of magnitude                ##
###########################################################################


set.seed(93)
cost <- exp(runif(p, 1, 10))

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 500, trace = TRUE, ofgmax=ofgmax)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))
###########################################################################
##                random cost different order of magnitude                ##
###########################################################################

# set.seed(93)
# cost <- exp(runif(p, 1, 10))

# clarsP <- clars(xtrain, ytrain, cost, maxk = 100, trace = TRUE)
# larsP <- lars(xtrain, ytrain, type = "lar")
# lassP <- lars(xtrain, ytrain, type = "lasso")

# clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
# larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
# lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

# xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
#           max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
# ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
#           max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

# plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


# clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
# larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
# lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

# xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
#           max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
# ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
#           max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

# plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

```

\newpage

Example 4: best cost more
=========================

```{r bestmore}

par(mfrow = c(1, 2))
###########################################################################
##                best cost more (best = largest ols coef                ##
###########################################################################

set.seed(95)
cost <- sort(runif(p, 1, 10))[best]

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 500, trace = FALSE)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

###########################################################################
##                best cost more (best = largest ols coef                ##
###########################################################################

# set.seed(95)
# cost <- sort(runif(p, 1, 10))[best]

# clarsP <- clars(xtrain, ytrain, cost, maxk = 500, trace = FALSE)
# larsP <- lars(xtrain, ytrain, type = "lar")
# lassP <- lars(xtrain, ytrain, type = "lasso")

# clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
# larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
# lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

# xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
#           max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
# ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
#           max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

# plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


# clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
# larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
# lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

# xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
#           max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
# ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
#           max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

# plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

```

\newpage

Example 5: best cost less
=========================

```{r bestless}
par(mfrow = c(1, 2))
###########################################################################
##               best cost less (best = largest ols coef)                ##
###########################################################################

set.seed(95)
cost <- sort(runif(p, 1, 10))[worst]

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 500, trace = FALSE)
mlasso(xtrain ,ytrain)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))
###########################################################################
##               best cost less (best = largest ols coef)                ##
###########################################################################

# set.seed(95)
# cost <- sort(runif(p, 1, 10))[worst]

# clarsP <- clars(xtrain, ytrain, cost, maxk = 500, trace = FALSE)
# mlasso(xtrain ,ytrain)
# larsP <- lars(xtrain, ytrain, type = "lar")
# lassP <- lars(xtrain, ytrain, type = "lasso")

# clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
# larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
# lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

# xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
#           max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
# ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
#           max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

# plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


# clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
# larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
# lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

# xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
#           max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
# ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
#           max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

# plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))
```

\newpage

Example 6: non-additive costs
=============================

If a blood measurement is already included, each subsequent blood measurement
is "2" less than it would otherwise be.

```{r nonadd}

par(mfrow = c(1, 2))

###########################################################################
##                  random cost same order of magnitude                  ##
###########################################################################

set.seed(95)
cost <- runif(p, 1, 10)
costfunc <- function(x)
{
    n <- sum(x[4:10])
    if (n > 1)
    {
        return(sum(cost[x]) - n * 2)
    } else {
        return(sum(cost[x]))
    }
}


clarsP <- clarstree(xtrain, ytrain, cost, maxk = 100, trace = FALSE, costfunc = costfunc)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


###########################################################################
##                  random cost same order of magnitude                  ##
###########################################################################

# set.seed(95)
# cost <- runif(p, 1, 10)
# costfunc <- function(x)
# {
#     n <- sum(x[4:10])
#     if (n > 1)
#     {
#         return(sum(cost[x]) - n * 2)
#     } else {
#         return(sum(cost[x]))
#     }
# }


# clarsP <- clars(xtrain, ytrain, cost, maxk = 100, trace = FALSE, costfunc = costfunc)
# larsP <- lars(xtrain, ytrain, type = "lar")
# lassP <- lars(xtrain, ytrain, type = "lasso")

# clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
# larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
# lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

# xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
#           max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
# ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
#           max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

# plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))


# clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
# larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
# lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

# xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
#           max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
# ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
#           max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

# plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
#      col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

# points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
# points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

# legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

```

\newpage

Example 7:  half cost worse version
===================================

Duplicate variables, add random noise, noisy variables cost half the
"originals"

```{r halfcostextra}

par(mfrow = c(1, 2))

# TODO: better strategy: normalize by saturated model cost?
library(lars)
library(cosso)
data(diabetes)
x <- scale(diabetes$x)
n <- nrow(x)
p <- ncol(x)
x2e <- matrix(rnorm(n * p, 0, 0.5), nrow = n, ncol = p)
x2 <- x + x2e
set.seed(95)
idx <- sample(1:n, round(n * 0.2), replace = FALSE)
x2[ , 2] <- x[ , 2]
x2[idx, 2] <- ifelse(x[idx, 2] == unique(x[ , 2])[1], unique(x[ , 2])[2],
                     unique(x[ , 2])[1])
x <- cbind(x, x2)
x <- scale(x)
p <- ncol(x)

y <- diabetes$y
maxk <- 50
eps <- 1e-9
best <- order(abs(lm(y ~ x - 1)$coef))
worst <- order(-abs(lm(y ~ x - 1)$coef))

set.seed(95)
train <- sample(1:n, round(n * 0.8), replace = FALSE)
test <- setdiff(1:n, train)
xtrain <- x[train, ]
ytrain <- y[train]
xtrain <- scale(xtrain)
ytrain <- ytrain - mean(ytrain)
xtest <- x[test, ]
ytest <- y[test]


###########################################################################
##                         half cost duplicates                          ##
###########################################################################


set.seed(95)
cost <- runif(p/2, 1, 10)
cost <- c(cost, cost / 2)

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 500, trace = FALSE)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

```

\newpage

Example 8: double cost worse version
====================================

Duplicate variables, add random noise, noisy variables cost twice the
"originals"

```{r doublecostextra}

par(mfrow = c(1, 2))

###########################################################################
##                        double cost duplicates                         ##
###########################################################################

set.seed(95)
cost <- runif(p/2, 1, 10)
cost <- c(cost, cost * 2)

clarsP <- clarstree(xtrain, ytrain, cost, maxk = 500, trace = FALSE)
larsP <- lars(xtrain, ytrain, type = "lar")
lassP <- lars(xtrain, ytrain, type = "lasso")

clarsP.eval <- evalclars(clarsP, xtrain, ytrain, cost)
larsP.eval <- evalclars(list(beta = larsP$beta), xtrain, ytrain, cost)
lassP.eval <- evalclars(list(beta = lassP$beta), xtrain, ytrain, cost)

xlim <- c(min(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost),
          max(clarsP.eval$modelcost, larsP.eval$modelcost, lassP.eval$modelcost))
ylim <- c(min(clarsP.eval$score, larsP.eval$score, lassP.eval$score),
          max(clarsP.eval$score, larsP.eval$score, lassP.eval$score))

plot(clarsP.eval$score ~ clarsP.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsP.eval$score ~ larsP.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassP.eval$score ~ lassP.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

clarsT.eval <- evalclars(clarsP, xtest, ytest, cost)
larsT.eval <- evalclars(list(beta = larsP$beta), xtest, ytest, cost)
lassT.eval <- evalclars(list(beta = lassP$beta), xtest, ytest, cost)

xlim <- c(min(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost),
          max(clarsT.eval$modelcost, larsT.eval$modelcost, lassT.eval$modelcost))
ylim <- c(min(clarsT.eval$score, larsT.eval$score, lassT.eval$score),
          max(clarsT.eval$score, larsT.eval$score, lassT.eval$score))

plot(clarsT.eval$score ~ clarsT.eval$modelcost, type = "p", xlim = xlim, ylim = ylim,
     col = "blue", pch = 16, cex = 4, xlab = "Model Cost", ylab = "Score")

points(larsT.eval$score ~ larsT.eval$modelcost, col = "red", pch = 16, cex = 3)
points(lassT.eval$score ~ lassT.eval$modelcost, col = "purple", pch = 16, cex = 2)

legend('topright', c("clars", "lars", "lasso"), pch = 16, col = c("blue", "red", "purple"))

```
